Loay Jabre; Last updated: 2023-04-27
- These scripts estimate how much of the pMetal we can account for using metaP data

Load packages - 2023-03-08
```{r}
#plotting 
library(ggplot2)
library(ggplotify)
library(cowplot)
library(ggridges)
library(ggpubr)
library(gridExtra)
library(grid)
library(gplots)
library(plotly)
library(ggthemes)
library(ggh4x) #mapping multiple aesthetics to colors 
library(patchwork) #allows for plot insets. e.g. map and the DFe and DMn in Figure 1
library (ggrepel) #makes point labels not touch each other in RA plots
#library(treemapify)
 
#other
library(data.table)
library(dplyr)
library(magrittr)
library(reshape2)
library(stringr)
library(scales)
library(tidyr)
library(tidyverse)

#colorblind friendly colors 
#https://rdrr.io/cran/ggthemes/man/colorblind.html

#list of colors 
#https://www.computerhope.com/htmcolor.htm #colors

#multi-pannel plots
#http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/
```
Import data of interest - 2023-03-30
```{r}
setwd("D:/School/PhD/PS117/data/") #this is the directory where I keep all the data

#import non-protein data
nonproteindata <- read.csv("LJ_ps117_allnonproteindata_20230308.csv", header = T)
nonproteindata_long <- melt (nonproteindata, id.vars = c(1:15, 17))

#import protein data
proteindata_unnorm <- read.csv("all_ps117_taxon-function_peptides_non-normalized_injections-means_20230324.csv", header = T)

#arrange the 'uniq_cluster' annotations alphabetically in each row to reduce redundancies. 
proteindata_unnorm$uniq_cluster <- sapply(lapply(strsplit(proteindata_unnorm$uniq_cluster, split = ",\\s*"), sort), paste, collapse = ", ") 

### a quick test to show how to normalize to all of proteome, vs to group
# https://stackoverflow.com/questions/50332613/normalize-by-group-for-all-columns
#group <- c("A", "A", "B", "B", "C", "C", "C", "A")
#abundance <- c(1,3,4,5,6,7,8,0)
#all_norm <- c(1,3,4,5,6,7,8,0)
#grp_norm <- c(1,3,4,5,6,7,8,0)
#x <- data.frame(group, abundance, all_norm, grp_norm) %>% mutate_at (3, funs ((. / sum(.)))) # %>%   group_by(group) %>% #mutate_at (4, funs ((. / sum(.))
```

Calculate theoretical maximum possible amount of Cd-CA that can be made 
```{r}
#total moles of Cd-CA, assuming 1mole Cd  = 1mol Cd-CA
ggplot(filter (nonproteindata, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes(x = PCd_nM, 
        y = PCd_nM *1 ,
        color = tempfe_treatment, 
        shape = tempfe_treatment))+
    geom_point(size = 4, alpha =0.65, stroke =1.5)+
    scale_color_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
   scale_shape_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
   labs (x = "PCd (nM)", y = "Theoretical maximum Cd-CA (nM) " )+
   facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))




#Convert moles of Cd-CA to ug. Assuming One mole Cd-CA = 68.76 kg
#1 mole Cd-CA = 68760 g  = 6.876 e+10 ug
 #– This is calculated based on Cd-CA AA sequence from Lane et al. 2005 (Cd-CA = 68.76 kDa = 68.76 kg mole -1)
#1 nmole Cd-CA = 6.876 e+1 ug


nonproteindata$theor_CdCA_ug <- nonproteindata$PCd_nM * 68.76

ggplot(filter (nonproteindata, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes(x = PCd_nM, 
        y = theor_CdCA_ug ,
        color = tempfe_treatment, 
        shape = tempfe_treatment))+
    geom_point(size = 4, alpha =0.65, stroke =1.5)+
    scale_color_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
   scale_shape_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
   labs (x = "PCd (nM)", y = "Theoretical maximum Cd-CA (ug/L) " )+
   facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))
```

Is the theoretical maximum reasonable? 
```{r}
ggplot(filter (nonproteindata, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes(#x = protein_ug_L*3, #this is to adjust for the wrong protein calculation - need to go back and double check all of this
          x = PON_ug_L*4.78,
          y = theor_CdCA_ug, 
          color = tempfe_treatment, 
         shape = tempfe_treatment))+
      scale_color_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
   scale_shape_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
       
    geom_point(size = 4, alpha =0.65, stroke =1.5)+
   labs (x = "Total protein estimate (ug/L) PON*4.78", y = "Theoretical maximum Cd-CA (ug/L) " )+
   facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))  
  # xlim (0,1200) + 
  # ylim (0,1200)




#Fraction of total protein that could be Cd-Ca

ggplot(filter (nonproteindata, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes(#x = protein_ug_L*3, #this is to adjust for the wrong protein calculation - need to go back and double check all of this
          y = 100*(theor_CdCA_ug/(PON_ug_L*4.78)),
           #y=(theor_CdCA_ug/(protein_ug_L)),
          x = bioassay))+
    geom_boxplot() + 
   geom_point(size = 3, alpha =0.65, stroke =1.5, aes (color = tempfe_treatment, 
         shape = tempfe_treatment))+

      scale_color_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
      scale_shape_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
       
   labs (x = "", y = "Maximum contribution of Cd-CA to total protein (%)" )+
  # facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))  

```

Look at Cd-CA 
```{r}
norm_proteomicsdata <- proteindata_unnorm  %>% mutate_at (40:69, funs ((. / sum(.)))) #this normalizes the data to total peptide abundance

cdca <- filter(norm_proteomicsdata, grepl("clust_6573", uniq_cluster))
cdca2 <- cdca [c(4, 40:69)]

cdca3 <- melt(cdca2, id.vars=c("uniq_cluster")) %>%
            rename(treatment = variable) %>%
            rename(norm_abundance = value ) %>% 
            separate(treatment, into =  c("bioassay", "timepoint", "temperature_treatment", "iron_treatment", "replicate"))  
 
cdca3$tempfe_treatment <- paste (cdca3$temperature_treatment, cdca3$iron_treatment, sep = "_")
  
cdca4 <- aggregate(cdca3$norm_abundance, by=list(cdca3$bioassay, cdca3$tempfe_treatment, cdca3$replicate), FUN = sum) %>%
          rename (bioassay = Group.1, 
          tempfe_treatment = Group.2, 
          replicate = Group.3, 
          cdca_norm_abundance = x)

ggplot(cdca4, aes(x = bioassay, 
                  y = 100*(cdca_norm_abundance)))+
    geom_boxplot() + 
    geom_point(size = 3, alpha =0.65, stroke =1.5, aes (color = tempfe_treatment, 
         shape = tempfe_treatment))+

      scale_color_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
      scale_shape_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
       
   labs (x = "", y = "Contribution of Cd-CA to total matched protein abundance (%)" )+
  # facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))  

cdca4$tempfe_treatment <- ifelse (cdca4$tempfe_treatment ==  "T0_T0", "T0",
                                                 cdca4$tempfe_treatment)

alldata2 <- merge (nonproteindata, cdca4, by = c("bioassay", "tempfe_treatment", "replicate"), all=TRUE)


#I was thinking just to calculate the proportion of the metal that can be accounted for by protein.

#step 1: Calculate moles of pCd from mass of Cd-CA
#each kg of Cd-CA = 0.0145 moles

#1 ug of cdca = 1.454 e-11 moles

#calculate ug/L of cdca: cdca_norm_abundance*PON_ug_L*4.78)
#calculate moles/L of cdca from ug/L : 1.454e-11 *(cdca_norm_abundance*PON_ug_L*4.78)

#calculate nM pCd from moles/L cd-ca 1e9 *(1.454e-11 *(cdca_norm_abundance*PON_ug_L*4.78))
ggplot(filter (alldata2, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes( x = bioassay, 
           #y = (cdca_norm_abundance*PON_ug_L*4.78))+
           y =(1e9 *(1.454e-11 *(cdca_norm_abundance*PON_ug_L*4.78)))/PCd_nM*100))+
            #y = (1e10 *(1.454e-11*(cdca_norm_abundance*PON_ug_L*4.78)))))+

   geom_boxplot() + 
    geom_point(size = 3, alpha =0.65, stroke =1.5, aes (color = tempfe_treatment, 
         shape = tempfe_treatment))+

      scale_color_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
      scale_shape_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
      labs (x = "", y = "Percent of pCd accounted for by Cd-CA" )+
  # facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))  
```

Look at Mn cluster
```{r}
#clust_5856
#clust_1906 - psbO
#clust_2763 - psbQ
oec <- filter(norm_proteomicsdata, grepl("clust_1906", uniq_cluster))

oec2 <- oec [c(4, 40:69)]

oec3 <- melt(oec2, id.vars=c("uniq_cluster")) %>%
            rename(treatment = variable) %>%
            rename(norm_abundance = value ) %>% 
            separate(treatment, into =  c("bioassay", "timepoint", "temperature_treatment", "iron_treatment", "replicate"))  
 
oec3$tempfe_treatment <- paste (oec3$temperature_treatment, oec3$iron_treatment, sep = "_")
  
oec4 <- aggregate(oec3$norm_abundance, by=list(oec3$bioassay, oec3$tempfe_treatment, oec3$replicate, oec3$uniq_cluster), FUN = sum) %>%
          rename (bioassay = Group.1, 
          tempfe_treatment = Group.2, 
          replicate = Group.3, 
          cluster = Group.4, 
          oec_norm_abundance = x)

oec4$normabunancadjusted <- ifelse (oec4$cluster ==  "clust_1906", oec4$oec_norm_abundance*3.0303e-11, 
                                    ifelse(oec4$cluster == "clust_2763", oec4$oec_norm_abundance*6.6667e-11, 
                                           ifelse(oec4$cluster == "clust_5856", oec4$oec_norm_abundance*4.347e-11, 
                                         "else")))
                                    
oec4$tempfe_treatment <- ifelse (oec4$tempfe_treatment ==  "T0_T0", "T0",
                                                 oec4$tempfe_treatment)

alldata2 <- merge (nonproteindata, oec4, by = c("bioassay", "tempfe_treatment", "replicate"), all=TRUE)


#I was thinking just to calculate the proportion of the metal that can be accounted for by protein.
#Calculate moles of pMn from mass of Mn-cluster

#each kg of OEC-cluster = 0.0303moles

#1 ug of oec = 3.0303 e-11 moles

#calculate ug/L of oec: oec_norm_abundance*PON_ug_L*4.78)
#calculate moles/L of oec from ug/L : 3.0303e-11 *(oec_norm_abundance*PON_ug_L*4.78)

#calculate nM pMn from moles/L cd-ca 1e9 *(3.0303e-11 *(oec_norm_abundance*PON_ug_L*4.78))





ggplot(filter (alldata2, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes( x = bioassay, 
           #y = (cdca_norm_abundance*PON_ug_L*4.78))+
          y = (1e9* as.numeric(normabunancadjusted)*PON_ug_L*4.78 )/PMn_nM*100))+
            #y = (1e10 *(1.454e-11*(cdca_norm_abundance*PON_ug_L*4.78)))))+

   geom_boxplot() + 
    geom_point(size = 3, alpha =0.65, stroke =1.5, aes (color = tempfe_treatment, 
         shape = tempfe_treatment))+

      scale_color_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
      scale_shape_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
      labs (x = "", y = "Percent of Mn accounted for by Mn-cluster protein" )+
  # facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))  



```

Mn cluster
```{r}

#clust_1906 - psbO
#clust_2763 - psbQ
oec <- filter(norm_proteomicsdata, grepl("clust_1906", cluster))
oec2 <- oec [c(3, 24:53)]

oec3 <- melt(oec2, id.vars=c("cluster")) %>%
            rename(treatment = variable) %>%
            rename(norm_abundance = value ) %>% 
            separate(treatment, into =  c("bioassay", "timepoint", "temperature_treatment", "iron_treatment", "replicate"))  
 
oec3$tempFe <- paste (oec3$temperature_treatment, oec3$iron_treatment, sep = "_")
  

oec4 <- aggregate(oec3$norm_abundance, by=list(oec3$bioassay, oec3$tempFe, oec3$replicate), FUN = sum) %>%
          rename (bioassay = Group.1, 
          tempFe = Group.2, 
          replicate = Group.3, 
          oec_norm_abundance = x)

alldata2 <- merge (alldata, oec4, by = c("bioassay", "tempFe", "replicate"), all=TRUE)

#I was thinking just to calculate the proportion of the metal that can be accounted for by protein.
#Calculate moles of pMn from mass of Mn-cluster

#each kg of OEC-cluster = 0.0303moles

#1 ug of oec = 3.0303 e-11 moles

#calculate ug/L of oec: oec_norm_abundance*PON_ug_L*4.78)
#calculate moles/L of oec from ug/L : 3.0303e-11 *(oec_norm_abundance*PON_ug_L*4.78)

#calculate nM pMn from moles/L cd-ca 1e9 *(3.0303e-11 *(oec_norm_abundance*PON_ug_L*4.78))

ggplot(filter (alldata2, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes(#x=PCd_nM,
        x = factor(tempFe, level = tempFe_order),
       #x = bioassay, 
       #y = (PCd_nM*68.76)/(PON_ug_L*4.78),
         y = (4 * 1e9 * 3.0303e-11 * oec_norm_abundance * PON_ug_L * 4.78)/PMn_nM*100,
       color = tempFe, 
       shape = tempFe))+
  #geom_point(size = 4, alpha =0.2)+
   stat_summary(fun = mean, geom = "point", size = 4)+
  stat_summary(fun.data =  mean_se, geom = "errorbar", show.legend = F, size = 1, width = 0.1)+
    scale_color_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      values = c('black', 'black', 'black','darkorange2', "darkorange2")) +
  
     scale_shape_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      values = c(5, 1, 19,1, 19)) +
  labs (x = "", y = "Percent of pMn accounted for by Mn-cluster" )+
    theme_bw()+
       facet_wrap(~bioassay) +
  theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))
```

look at Rubisco
```{r}
rubisco <- filter(norm_proteomicsdata, grepl("^clust_2490$|^clust_129$", uniq_cluster)) #clust_2490| clust_129
rubisco <- rubisco [c(4, 40:69)]

rubisco2 <- melt(rubisco, id.vars=c("uniq_cluster")) %>%
            rename(treatment = variable) %>%
            rename(norm_abundance = value ) %>% 
            separate(treatment, into =  c("bioassay", "timepoint", "temperature_treatment", "iron_treatment", "replicate"))  
 
rubisco2$tempfe_treatment <- paste (rubisco2$temperature_treatment, rubisco2$iron_treatment, sep = "_")
  
rubisco3 <- aggregate(rubisco2$norm_abundance, by=list(rubisco2$bioassay, rubisco2$tempfe_treatment, rubisco2$replicate), FUN = sum) %>%
          rename (bioassay = Group.1, 
          tempfe_treatment = Group.2, 
          replicate = Group.3, 
          rubisco_norm_abundance = x)





ggplot(rubisco3, aes(x = tempfe_treatment, 
                  y = 100*(rubisco_norm_abundance)))+
   # geom_boxplot() + 
    geom_point(size = 3, alpha =0.65, stroke =1.5, aes (color = tempfe_treatment, 
         shape = tempfe_treatment))+

      scale_color_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
      scale_shape_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT -Fe", "LT +Fe", "HT -Fe", "HT +Fe"),
                      values = c(5, 1, 19,1, 19)) +
       
   labs (x = "", y = "Contribution of Rubisco to total matched protein abundance (%)" )+
   facet_wrap(~bioassay) +
   theme_bw()+
   theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))  



```

Protein per L and prelimnary plots to make sure data are good
```{r}
plot1 <- function (xaxis, yaxis, xaxislabel, yaxislabel) {ggplot(filter (nonproteindata, grepl ("T0|T8|", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes(x = {{xaxis}}, 
           y = {{yaxis}},
           color = tempfe_treatment, 
           shape = tempfe_treatment))+
       geom_point(size = 4, alpha =0.65, stroke =1.5)+
       scale_color_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      values = c('purple', 'black', 'black','darkorange2', "darkorange2")) +
       scale_shape_manual(name="",
                      breaks = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      values = c(5, 1, 19,1, 19)) +
       labs (x = xaxislabel, y = yaxislabel)+
       theme_bw()+
       facet_wrap(~bioassay) +
       theme(axis.title = element_text (size = 14, face = "bold"), 
             axis.text = element_text(size = 12, color = 'black'), 
             legend.title = element_text(size = 14, face = "bold"),
             legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
             strip.background =element_rect(fill = "white"),
             strip.text.x = element_text(size = 14, face = "bold"))
}

plot1 (protein_ug_L, total_chla_ug_L, expression ("Total protein"~(mu*g~L^-1)), expression ("Total chl-a"~(mu*g~L^-1)))

plot1 (protein_ug_L, PON_ug_L, expression ("Total protein"~(mu*g~L^-1)), expression ("PON"~(mu*g~L^-1)))

plot1 (protein_ug_L, POC_ug_L, expression ("Total protein"~(mu*g~L^-1)), expression ("POC"~(mu*g~L^-1)))

plot1 (POC_ug_L, total_chla_ug_L, expression ("PON"~(mu*g~L^-1)), expression ("Total chl-a"~(mu*g~L^-1)))


         #x=iron_treatment, 
       #y = total_chla_ug_L,
       #y = protein_ug_per_L,
       #x = PON_ug_L,
       #x = PON_ug_L*4.78,

       #y = (PCd_nM*68.76),
          #y = (PCd_nM),
       #x= PCd_nM/PON_uM,






```

Remember to use total protein feature abundance instead of ID'd peptides only
```{r}

```

SOD cluster
```{r}
sod <- filter(norm_proteomicsdata, grepl("^clust_2501$", cluster))
sod2 <- sod [c(3, 24:53)]

sod3 <- melt(sod2, id.vars=c("cluster")) %>%
            rename(treatment = variable) %>%
            rename(norm_abundance = value ) %>% 
            separate(treatment, into =  c("bioassay", "timepoint", "temperature_treatment", "iron_treatment", "replicate"))  
 
sod3$tempFe <- paste (sod3$temperature_treatment, sod3$iron_treatment, sep = "_")

sod4 <- aggregate(sod3$norm_abundance, by=list(sod3$bioassay, sod3$tempFe, sod3$replicate), FUN = sum) %>%
          rename (bioassay = Group.1, 
          tempFe = Group.2, 
          replicate = Group.3, 
          sod_norm_abundance = x)

alldata2 <- merge (alldata, sod4, by = c("bioassay", "tempFe", "replicate"), all=TRUE)


#I was thinking just to calculate the proportion of the metal that can be accounted for by protein.
#Calculate moles of pMn from mass of Mn-cluster

#each kg of sod-cluster = 0.0303moles

#1 ug of sod = 3.0303 e-11 moles

#calculate ug/L of sod: sod_norm_abundance*PON_ug_L*4.78)
#calculate moles/L of sod from ug/L : 3.0303e-11 *(sod_norm_abundance*PON_ug_L*4.78)

#calculate nM pMn from moles/L cd-ca 1e9 *(3.0303e-11 *(sod_norm_abundance*PON_ug_L*4.78))

ggplot(filter (alldata2, grepl ("T0|T8", full_treatment_name) & grepl ("BA2|", bioassay)), 
       aes(#x=PCd_nM,
       #x = PCd_nM,
       x = factor(tempFe, level = tempFe_order),
       #x = bioassay, 
       #y = (PCd_nM*68.76)/(PON_ug_L*4.78),
         y = (1e9 *(4e-11 *(sod_norm_abundance*PON_ug_L*4.78)))/PMn_nM*100,
       color = tempFe, 
       shape = tempFe))+
  geom_point(size = 4, alpha =0.2)+
   stat_summary(fun = mean, geom = "point", size = 4)+
  stat_summary(fun.data =  mean_se, geom = "errorbar", show.legend = F, size = 1, width = 0.1)+
    scale_color_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      values = c('black', 'black', 'black','darkorange2', "darkorange2")) +
  
     scale_shape_manual(name="",
                      breaks = c("T0_T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      labels = c("T0", "LT_noFe", "LT_Fe", "HT_noFe", "HT_Fe"),
                      values = c(5, 1, 19,1, 19)) +
  labs (x = "", y = "Percent of pMn accounted for by Fe-Mn SOD" )+
    theme_bw()+
       facet_wrap(~bioassay) +
  theme(axis.title = element_text (size = 14, face = "bold"), 
        axis.text = element_text(size = 12, color = 'black'), 
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 10, margin = margin(r = 0.5, unit = 'cm')), 
        strip.background =element_rect(fill = "white"),
        strip.text.x = element_text(size = 14, face = "bold"))
```